---
title: "Announcing Rosa"
output: 
  rmarkdown::github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(babynames)
library(factoextra)
library(gghighlight)
library(ggthemes)
library(waffle)
library(png)
library(grid)
library(gridExtra)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

## Add image to plots
img <- readPNG("rosa2.png")
```

A quick announcement that Rosa was born on August 7, 2020 at 9:39 AM.

Family and friends: see our [Baby Instagram](https://www.instagram.com/bamacchetti/) for pictures.

Now for the numbers and context.

## Her stats

First of all, her basic statistics. These are the first questions almost everyone has about babies - how heavy and how long? It's helpful to put these numbers in context by looking at how Rosa compares with other Caucasian babies. 

At **3.3 kilos (7 lbs 5 oz)** Rosa sits comfortably around the average. If all babies were exactly the same size, we wouldn't have a distribution here (the "bell curve" below). We probably wouldn't have that many types of sports either if humans had the same body type.

I used data from [this article](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2802014/) by Janssen et al. (2017).

```{r weight}
ggplot(data.frame(x = c(1,6)), aes(x)) +
    # plotting without data
  stat_function(fun = dnorm, args = list(mean = 3.455, sd = .434),
                size = 2, alpha = 0.3) +
    # Add normal curve here - babies in BC at 39 weeks
  geom_vline(xintercept = 3.3, color = "#CC6677", size = 2) +
  theme_fivethirtyeight() +
  labs(title = "Slightly small but full of energy", # "Weighing in at 3.3 kilograms (7 lbs 5 oz)",
       subtitle = "A bit below average but full of energy",
       x = "Weight in Kilograms") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  scale_x_continuous(n.breaks = 6) +
  annotation_custom(rasterGrob(img),
                    xmin = 2, xmax = 3, ymin = .3, ymax = 1) +
  annotate("text", x = 2.5, y = .85, 
           label = "Rose clocks in at\n 3.3. kilos (7 lbs 5 oz)",
           color = "#CC6677")
```

Length wise, though, Rosa is somewhat above the average.

```{r length}
ggplot(data.frame(x = c(40, 60)), aes(x)) +
  # plotting without data
  stat_function(fun = dnorm, args = list(mean = 49.4, sd = 1.7), 
                size = 2, alpha = 0.3) +
  # Add normal curve here - babies in BC at 39 weeks
  geom_vline(xintercept = 51.3, color = "#CC6677", size = 2) +
  theme_fivethirtyeight() +
  labs(title = "Longer than average",
       subtitle = "Above average, maybe not a basketball player",
       x = "Length in centimeters") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  scale_x_continuous(n.breaks = 6) +
  annotation_custom(rasterGrob(img),
                    xmin = 50, xmax = 60, ymin = .1, ymax = .175) +
  annotate("text", x = 55, y = .2, 
           label = 'Rosa was\n 51.3 cm (1 ft 8 in)',
           color = "#CC6677")
```

## What's in a name?

Rosa was named after her great grandmother. Her name follows the Portuguese pronounciation, which differs mainly in the "R" - it's different than in English or the rolling R in Spanish. It's tricky. See the pronunciation of Rosa [here](https://www.infopedia.pt/dicionarios/lingua-portuguesa/rosa) and a breakdown of the "R" sound [here](https://youtu.be/3jdovKE2f4I?t=366).

## How common is Rosa as a name?

In the United States in 2017, around 400 Rosas were born, which is 0.02% of all children that year. 

The big question is: **What is the possibility that Rosa will be in a class with someone with the same name?** And have to go by Rosa S. and always get confused with another kid?

It turns out the chance Rosa will be in a class with another Rosa is about **1 in 200**. Which is pretty unlikely. 

Here I'm assuming classes are 20 kids and the popularity of names in 2020 is similar to 2017.

```{r common name}
rosa <- c(1, 199)

waffle(rosa, rows = 10,
       colors = c("#CC6677", "lightgray")) +
  labs(title = "Rosa is a fairly unique name in the US",
       subtitle = "1 in 200 chances she'll be in a class with another Rosa") +
  theme_fivethirtyeight() +
  theme(axis.text = element_blank(),
        legend.position = "none",
        plot.background = element_rect(fill = "white"))
```

Unfortunately for girls named Emma, the most popular name in 2017 (and 1% of all babies), their chances are 1 in 5.

```{r common name 2, fig.width=4, fig.height=2}
emma <- c(1, 4)

waffle(emma, rows = 1,
       colors = c("#44AA99", "lightgray")) +
  labs(title = "Emma is a very common name",
       subtitle = "1 in 5 chances") +
  theme_fivethirtyeight() +
  theme(axis.text = element_blank(),
        legend.position = "none",
        plot.background = element_rect(fill = "white"))
```

## How popular was the name Rosa over time?

Rosa is a pretty classic name, going at least back to 1880, when the Social Security Administration started tracking baby names.

It peaked in popularity in two decades: the 1920s and 1950s. Perhaps the influence of Rosa Luxemburg and Rosa Parks?

```{r name, fig.width=8}
#head(babynames)

babynames %>% filter(name == "Rosa", sex == "F") %>%
  ggplot(aes(year, n)) +
    geom_line(color = "#CC6677", size = 2) +
    #geom_smooth(span = 0.2, se = FALSE, color = "red") +
    theme_fivethirtyeight() +
    labs(title = "Rosa as a name was most popular in the 1920s and 1950s",
         subtitle = "Source: United States Social Security Administration",
         x = "Year", y = "Number of baby girls") +
    annotation_custom(rasterGrob(img),
                      xmin = 1970, xmax = 1995, ymin = 1500, ymax = 1900)
```

## What kind of names are similar?

**Given the popularity of Rosa over time, are there names with a similar trend?** It seems Miriam is a close contender, along with Priscilla. Jenny (as opposed to Jennifer) had a similar trend, but peaked in the 1970s. Cassidy is an example of a name that is relatively new, but like the other names, is less popular since the 2000s.

Thanks to [this analysis](https://rpubs.com/jalapic/babynames) for the inspiration and the code.

```{r name2}
# Pivot to do a PCA
# https://rpubs.com/jalapic/babynames

babywide <- babynames %>%
  filter(sex == "F") %>%
  select(name, year, n) %>%
  pivot_wider(names_from = year, values_from = n, values_fill = 0) %>%
  column_to_rownames("name")

set.seed(123)
baby_kmeans <- kmeans(babywide, 9)

# check out names
#table(baby_kmeans$cluster)

# uncommon names get pulled into one large cluster
odd_names <- names(baby_kmeans$cluster[baby_kmeans$cluster == 2])

# or, find the cluster with Rosa in it
cluster_name <- names(baby_kmeans$cluster[baby_kmeans$cluster == 5])

# filter out uncommon names or limit to one cluster
babywide_common <- babywide %>%
  rownames_to_column("name") %>%
  #filter(!name %in% odd_names) %>%
  filter(name %in% cluster_name) %>%
  column_to_rownames("name")

# new kmeans
set.seed(123)
cbaby_kmeans <- kmeans(babywide_common, 6)

#table(cbaby_kmeans$cluster)

babylong <- data.frame(babywide_common, cbaby_kmeans$cluster) %>%
  rownames_to_column("name") %>%
  rename(cluster = cbaby_kmeans.cluster) %>%
  pivot_longer(-c(cluster, name), names_to = "year", values_to = "n") %>%
  mutate(year = as.numeric(str_remove(year, "X")))
  
#filter(babylong, name == "Rosa")

# sample 10 other names in the same cluster
set.seed(125)
other_names <- babylong %>% 
  filter(cluster == 5) %>% # insert cluster number here
  distinct(name) %>% 
  sample_n(10) %>%
  pluck("name")
```
```{r comparable names, fig.width=11, fig.height=7}
babylong %>%
  #filter(cluster == 2) %>%
  filter(name %in% other_names | name == "Rosa") %>%
  ggplot(aes(year, n, group = name, color = name)) +
    geom_line(size = 2) +
    gghighlight(name %in% c("Rosa", "Miriam", "Priscilla", "Jenny", "Cassidy"), 
                label_key = name, unhighlighted_params = list(alpha = 0)) +
    scale_y_log10() +
    theme_fivethirtyeight() +
    scale_color_ptol() +
    labs(title = "A few other names follow the same trend as Rosa",
         subtitle = "Unlike Cassidy, Rosa was consistently popular for the last century",
         x = "Year", y = "Number of baby girls")
```

## How the parents are doing

Inadvertently I had my smart watch on during the birth, which captured my heart rate over time. It's no surprise it went up after we saw the baby for the first time.

```{r hr, fig.width=9, fig.height=5}
hr <- read_csv("withings_data/raw_tracker_hr.csv")

hr2 <- hr %>%
  mutate(value2 = str_remove_all(value, "\\[|\\]") %>%
           str_split(",") %>% 
           map(1) %>%
           unlist() %>%
           as.numeric())

hr2 %>%
  filter(start > mdy_h("08-07-2020 0") & start < mdy_h("08-07-2020 24")) %>%
  ggplot(aes(start, value2)) +
    geom_line(aes(alpha = value2), color = "gray", size = 1.5) +
    geom_smooth(se = FALSE, linetype = 2, size = 1, color = "gray") +
    geom_vline(xintercept = mdy_hm("08-07-2020 09:39"),
               color = "#CC6677", size = 2) +
    annotation_custom(rasterGrob(img),
                     xmin = mdy_h("08-07-2020 10"), 
                     xmax = mdy_h("08-07-2020 14"), 
                     ymin = 110, ymax = 130) +
    theme_fivethirtyeight() +
    labs(title = "Father's heart rate before, during, and after Rosa's birth",
         subtitle = "Measured in beats per minute, with time of day in 24 hour time") +
    scale_x_datetime(date_labels = "%H", date_breaks = "2 hours")
```

My watch also tracks sleep, which gives a pretty accurate portrayal of pre and post-partum sleep habits. However, the watch can only record sleep that is longer than 3 hours per stint - the 9th and 10th were pretty tough.

```{r sleep}
sleep <- read_csv("withings_data/sleep.csv")

sleep2 <- sleep %>%
  mutate(date = as.Date(from),
         amount = as.numeric(difftime(to, from, "hours")))

sleep2 %>%
  filter(month(date) == 8,
         day(date) < 12) %>%
  ggplot(aes(date, amount, fill = amount)) +
    geom_col(color = "black") +
    theme_fivethirtyeight() +
    theme(legend.position = "none") +
    scale_y_continuous(breaks = seq(0, 10, by = 2),
                       ) +
    scale_fill_distiller(palette = "RdYlGn", direction = 1) +
    scale_x_date(date_breaks = "1 day", date_labels = "%d",
                 expand = c(0.01, 0)
                 ) +
    labs(title = "Father's nightly sleep in hours, during August",
         subtitle = "Can you guess the birth date?")
```

------

Welcome to the world, Rosa! There will be lots of adventures, numbers, and trends out there for you to explore.

Created by [Richard Bamattre](https://twitter.com/rbamattre). My partner says I have too much time on my hands since taking paternity leave.